Offline Mode for Corporate Mobile Client Security Architecture

### 2016-09-01 docs/papers/2016JGC/2016JGC.tex
adicao de prototipo de tela splash

\begin{abstract}
Preventing data leakage on the mobile app is a crucial security problem. Therefore, additional control and protection measures should be taken for the confidential data on the mobile devices that leave the boundaries of the organization. In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the app or of the business. This paper presents a novel approach to the security of the corporate mobile apps, in particular the offline mode. The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e. the protection of the confidential data in use when the mobile client is not connected to the corporate cloud. The protection of the sensitive data is provided by the combination of the cryptographic methods and analytics methods to detect malicious user behavior. The proposed security architecture supports the basic mobile app protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
\end{abstract}

@@

Additional security issues and requirements have to be considered when mobile apps are actively used in corporate cloud environment \cite{yovel2014}. Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile app is usually light-weighted and generally less protected. The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.

Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security. This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}. Moreover, our proposed approach includes the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

The paper is structured as follows. Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions related to the offline protection in the BYOD world. Section \ref{sec_the_client_security_architecture} outlines the basic processes and the module security infrastructure of the proposed solution. Section \ref{sec_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec_algorithms} provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS. Section 6 discusses the security proofs including the outline of adversary model, common threat scenarios and the proposed analytics to discover abnormal usage. Still in Section 6, complexity analysis and practical implementation results are included. Section 7 concludes the paper.

\section{Related works}
\label{sec_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the app at a maximum level with minimum resource requirements. The most commonly used are:

\begin{enumerate}
	\item Using geolocation of the device to trace its usage;
	\item Setting up the expiration period of an app;
	\item Setting up the expiration period of app pass pin;
	\item Setting up the counter of failed tries;
	\item Secured transfer politics between apps;
	\item Restricting access to the corporate app;
	\item Restricted or prohibited offline access;
	\item Logging and auditing.
\end{enumerate}

Preventing data leakage on the mobile device is a crucial security problem. Therefore, it is necessary to take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the app or the business itself?

From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud computing as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}. Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}. Obviously, it is natural the mobile app cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.

The necessity to use schemes that function without putting load on a provider arises when it is desired to make the app less dependent on the cloud provider, i.e. corporate app continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}.
Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.

In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure app wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access. To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the app and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The mobile security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile app security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The client app then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.

@@

\section{The proposed solution for offline mobile security}
\label{sec_offline_mode}
This section proposes an approach for the mobile app protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile app protection follow a model where the protected mobile app can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile app protection herein proposed are:

\begin{enumerate}
	\item Optimized communication with the cloud when the app does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
	\item Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
\end{enumerate}

The most important security issues in the proposed model arise when the app goes to the offline mode and the user is still allowed to get the access to the protected SME documents. In this case, the server can neither monitor the user activity nor provide the protection methods. The security should be performed at the mobile client. Additionally, the maximum protection should be provided at the minimum resource cost.

In the online mode the mobile app uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently. Thus, the proposal is that the authentication/authorization mechanisms in the offline mode should utilize the derived proof of the user identity. The requirements for the proof are as follows:

@@

Such proof cannot be stored on the mobile app, as it is not possible to guarantee its protection in the offline mode. Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the client and the user in a protected manner. In this case, there is no need to store the function of the user password. The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is difficult.

@@

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section \ref{sec_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

@@

The client app performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the client area. Note that the client app does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the app keeps track of the user activities and the key expiry. The complete list of the app activities is presented in the Figure \ref{fig:6}.

@@

The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section \ref{sec_algorithms}. The key feature of the offline security is that the client app does not actually store any part of the user password to be verified. The client app combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY. If the user provides the wrong share the client will not be able to recognize it, but will decrypt the files incorrectly. Additionally, the password entering is tracked and too many tries in a short time are considered a threat.

@@

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:8} depicts the Threat Intelligence Manager for offline behavioral analysis.

@@

As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the client app. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information of the captured operations to the User Activity Logging (UAL).

The UAL is responsible for the incremental logging, feature extraction and data modeling for malicious behavior detection, through the Log Analysis Center (LAC). The logged information shall be decomposed into selected features and modeled as matrices, composed of the number of occurrences of the selected features by its location and by the occurrence time. The resultant data is submited to the LAC, for anomaly detection.

@@

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile app is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:9}.

@@

The key hierarchy can be explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.

@@

\subsection{AES file encryption }
\label{sec_aes}
The protected files are encrypted on the server side with the secure 128/256 bit AES encryption, which is currently an industrial standard. Nevertheless, domain administrator can choose to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.

The symmetric encryption on a server side is performed in two steps. First there is a single AES key for preserving the server storage. Second, when the encrypted file is sent from the server to the mobile app or desktop device another unique AES key is generated. This key serves for the encryption on the client side. In other words, along with the encrypted file the user gets his own unique AES key for decrypting the files received. This key is protected by attribute-based public key encryption (ABE).

@@

The model supports the simple selective ABE scheme \cite{galibus2015cloud,galibus2014}. The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set $\{t_i\}_M$ is equal to the attribute in the set{ti}M, the corresponding user U can decrypt the text M. In other words, as soon as the user and share have one attribute in common – the user can get access to the share. The components of the ABE encryption are:

@@

Since the underlying sharing scheme is perfect \cite{galibus2008some}, hence the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.

The proposed authentication system is based on the shared storing of the user key. Also, the mobile app acts as a dealer in the SSS. Using the SSS ensures that the key can only be accessed by an authenticated user. The participants of the (2, 2)-threshold SSS are the user and device. The user share $s_1(x)$ is computed based on the PIN and the PASS entered by the user. Additionally, the current time value TIME is used in the calculation of the share. Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way function that transforms the data into the string of the desired length:

@@

Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data. Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.

MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the client app.

@@

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}},$, in order to obtain the vector of eigenvalues $e$, as following:

@@

\begin{enumerate}
	\item Steal the user keys one-by-one while the app performs decryption, if he is outsider;
	\item Use expired key to decrypt the files, if he is insider;
	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
	\item Steal the encrypted files but not the decrypted ones.
\end{enumerate}

@@

\textbf{Offline Mode}\\

In the offline mode the client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The only possible scenario for the information leakage in the case of the malicious outsider is if:

@@

The temporary nature of all parameters obliges the user to connect to server when necessary and prevents the malicious actions from the user side. The only possible scenario when malicious or expired user wishes to prolong his old credentials is:

\begin{enumerate}[label=(\alph*)]
	\item Steal the file and try to decrypt it with offline dictionary attacks.
	\item He has to be able to combine;
	\item He has to steal the KEY\_SET synchronized with his credentials;
	\item He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
	\item He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
\end{enumerate}

Basically, for a malicious user there is practically no way to use the client with the old keys. The fact that the client does not contain any data to be checked or validated does not prevent the user from seeing the contents of decrypted files, but with the wrong password the decrypted files will be different from the original ones.

The client still has to count the tries (to prevent the hacker to perform brute force attack) within one session. In accordance with the conditions 2 and 4 of the adversary model, the above actions are not possible for such adversary.
\\
\\
\textbf{Online mode}\\

In the online mode the protection of the mobile system is backed up by the communication with server. The key expiry period and the usage of J-PAKE \cite{hao2010j,hao2008password} guarantees that the sensitive data sent over the network cannot leak. The detailed security analysis of J-PAKE is presented in \cite{toorani2014security}. The data sent in the clear i.e. encrypted files and the public keys of ABE does not need to be protected additionally as the AES and ABE guarantee its security.

@@

Therefore, it is necessary to generate three matrices for malicious behaviors analysis, such as:

@@

\subsection{Complexity analysis }
\label{sec_complex}

>>>>

\begin{abstract}
Preventing data leakage on the mobile client is a crucial security problem. Therefore, additional control and protection should be taken for the confidential data on the mobile clients that leave the boundaries of the organization. In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the mobile client or of the business. This paper presents a novel approach to the security of the corporate mobile clients, in particular the offline mode. The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e. the protection of the confidential data in use when the mobile client is not connected to the corporate cloud. The protection of the sensitive data is provided by the combination of the cryptographic and analytics methods to detect malicious user behavior. The proposed security architecture supports the basic mobile client protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
\end{abstract}

@@

Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}. Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile client is usually light-weighted and generally less protected. The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.

Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the mobile client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile client has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security.

This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute} for securing mobile clients. Moreover, our proposed approach includes the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

The paper is structured as follows. Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions related to the offline protection in the BYOD world. Section \ref{sec_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec_algorithms} explains the algorithms for the core security methods: AES, ABE, SSS and MOS. Section \ref{sec_results} discusses the security proofs including the outline of adversary model, the common threat scenarios, the data modeling, the complexity analysis and dicussion of the practical implementation results. Section \ref{sec_conclusion} concludes the paper.

\section{Related works}
\label{sec_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the mobile client at a maximum level with minimum resource requirements. The most commonly used are:

\begin{enumerate}
	\item Using geolocation of the device to trace its usage;
	\item Setting up the expiration period of an app;
	\item Setting up the expiration period of client pass pin;
	\item Setting up the counter of failed tries;
	\item Secured transfer politics between apps;
	\item Restricting access to the corporate app;
	\item Restricted or prohibited offline access;
	\item Logging and auditing.
\end{enumerate}

Preventing data leakage on the mobile device is a crucial security problem. Therefore, it is necessary to take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the mobile client or the business itself?

From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}. Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}. Obviously, it is natural the mobile client cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.

The necessity to use schemes that function without putting load on a provider arises when it is desired to make the mobile client less dependent on the cloud provider, i.e. corporate client continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}.
Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.

In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure client wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access. To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the mobile client and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The mobile security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile client security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The mobile client then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.

@@

\section{The proposed solution for offline mobile security}
\label{sec_offline_mode}
This section proposes an approach for the mobile client protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile client protection follow a model where the protected client can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile client protection herein proposed are:

\begin{enumerate}
	\item Optimized communication with the cloud when the mobile client does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
	\item Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
\end{enumerate}

The most important security issues in the proposed model arise when the client goes to the offline mode and the user is still allowed to get the access to the protected SME documents. In this case, the server can neither monitor the user activity nor provide the protection methods. The security should be performed at the mobile client. Additionally, the maximum protection should be provided at the minimum resource cost.

In the online mode the mobile client uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently. Thus, the proposal is that the authentication/authorization mechanisms in the offline mode should utilize the derived proof of the user identity. The requirements for the proof are as follows:

@@

Such proof cannot be stored on the client, as it is not possible to guarantee its protection in the offline mode. Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the mobile client and the user in a protected manner. In this case, there is no need to store the function of the user password. The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is difficult.

@@

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

@@

The mobile client performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the mobile client area. Note that the mobile client does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the mobile client keeps track of the user activities and the key expiry. The complete list of the mobile client activities is presented in the Figure \ref{fig:6}.

@@

The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section \ref{sec_algorithms}. The key feature of the offline security is that the mobile client does not actually store any part of the user password to be verified. The mobile client combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY. If the user provides the wrong share the mobile client will not be able to recognize it, but will decrypt the files incorrectly. Additionally, the password entering is tracked and too many tries in a short time are considered a threat.

@@

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:8} depicts the TIM for offline behavioral analysis.

@@

As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

The UAL is responsible for the incremental logging, feature extraction and data modeling for malicious behavior detection, through the Log Analysis Center (LAC). The logged information shall be decomposed into selected features and modeled as matrices, composed of the number of occurrences of the selected features by its location and by time. The resultant data is submited to the LAC, for anomaly detection.

@@

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile client is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:9}.

@@

The key hierarchy is explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.

@@

\subsection{AES file encryption }
\label{sec_aes}
The protected files are encrypted on the server side with the secure 128/256 bit AES encryption, which is currently an industrial standard. Nevertheless, it is possible to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.

The symmetric encryption on a server side is performed in two steps. First there is a single AES key for preserving the server storage. Second, when the encrypted file is sent from the server to the mobile client another unique AES key is generated, which serves for the encryption on the client side. In other words, along with the encrypted file the user gets his own unique AES key for decrypting the files received. This key is protected by attribute-based public key encryption (ABE).

@@

The model supports the simple selective ABE scheme \cite{galibus2015cloud,galibus2014}. The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set $\{t_i\}_U$ is equal to the attribute in the set $\{t_i\}_M$, the corresponding user $U$ can decrypt the text $M$. In other words, as soon as the user and share have one attribute in common – the user can get access to the share. The components of the ABE encryption are:

@@

The adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.

The proposed authentication system is based on the shared storing of the user key. Also, the mobile client acts as a dealer in the SSS. Using the SSS ensures that the key can only be accessed by an authenticated user. The participants of the (2, 2)-threshold SSS are the user and device. The user share $s_1(x)$ is computed based on the PIN and the PASS entered by the user. Additionally, the current time value TIME is used in the calculation of the share. Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way function that transforms the data into the string of the desired length:

@@

Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data.

Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, which can be user operations, should be modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the mobile client.

@@

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $e$, as following:

@@

\begin{enumerate}
	\item Steal the user keys one-by-one while the mobile client performs decryption, if the attacker is outsider;
	\item Use expired key to decrypt the files, if the attacker is insider;
	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the mobile client;
	\item Steal the encrypted files but not the decrypted ones.
\end{enumerate}

@@

In the offline mode the mobile client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The possible scenario for the information leakage in the case of the malicious outsider is if:

@@

The temporary nature of all parameters obliges the user to connect to server when necessary and prevents the malicious actions from the user side. The possible scenario when malicious or expired user wishes to prolong his old credentials is:

\begin{enumerate}[label=(\alph*)]
	\item Steal the file and try to decrypt it with offline dictionary attacks.
	\item He has to be able to combine;
	\item He has to steal the KEY\_SET synchronized with his credentials;
	\item He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
	\item He has to do everything without the mobile client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
\end{enumerate}

Basically, for a malicious user there is practically no way to use the mobile client with the old keys. The fact that the client does not contain any data to be checked or validated does not prevent the user from seeing the contents of decrypted files, but with the wrong password the decrypted files will be different from the original ones.

The mobile client still has to count the tries (to prevent the hacker to perform brute force attack) within one session. In accordance with the conditions 2 and 4 of the adversary model, the above actions are not possible for such adversary.

@@

Therefore, it is necessary to generate three matrices for the following malicious behaviors analysis:

@@

\subsection{Complexity analysis}
\label{sec_complex}

The complexity analysis depends on the operation that the user and the mobile client perform in order to keep the app protected. In the offline mode the mobile client performs the following actions:
